---
title: "Spectrogram with av package"
author: "Jeroen Ooms"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Spectrogram with av package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(av)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First we create a 5 sec audio fragment to test with:


```{r}
# Get a shorter fragment
wonderland <- system.file('samples/Synapsis-Wonderland.mp3', package='av')
av_audio_convert(wonderland, 'short.mp3', total_time = 5)
```

Calculate the frequency data and plot the spectrogram:

```{r}
fft_data <- read_audio_fft('short.mp3')
plot(fft_data)
```

You can turn off dark mode to use the default R colors:

```{r}
plot(fft_data, dark = FALSE)
```

## Spectrogram video

You can also create a spectrogram video like this:

```r
av_audio_animation('short.mp3', width = 720, height = 480)
```

<video width="100%" controls>
<source src="https://jeroen.github.io/videos/spectrogram.mp4" type="video/mp4">
Your browser does not support the video tag.
</video>

## Compare with seewave

Compare spectrograms using the `tico` audio sample included with the seewave package:

```{r}
library(seewave)
library(ggplot2)
data(tico)
ggspectro(tico, ovlp = 50) + geom_tile(aes(fill = amplitude)) + stat_contour()
```

To use av, we first save the wav file and then create spectrogram:

```{r}
# First export wav file
savewav(tico, filename = 'tico.wav')
plot(read_audio_fft('tico.wav'))
```

## Compare with phonTools

Use the audio sample included with phonTools:

```{r}
library(phonTools)
data(sound)
spectrogram(sound, maxfreq = sound$fs/2)
```


Save the wav file and then create spectrogram. We match the default window function from phonTools:

```{r}
phonTools::writesound(sound, 'sound.wav')
plot(read_audio_fft('sound.wav', window = phonTools::windowfunc(1024, 'kaiser')))
```
