---
title: "Spectrograms in R using the 'av' package"
author: "Jeroen Ooms"
date: "`r Sys.Date()`"
---

```{r setup, include = FALSE}
library(av)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First we create a 5 sec audio fragment to test with:


```{r}
# Get a shorter fragment
wonderland <- system.file('samples/Synapsis-Wonderland.mp3', package='av')
av_audio_convert(wonderland, 'short.mp3', total_time = 5)
```

Calculate the frequency data and plot the spectrogram:

```{r}
fft_data <- read_audio_fft('short.mp3')
plot(fft_data)
```

You can turn off dark mode to use the default R colors:

```{r}
plot(fft_data, dark = FALSE)
```

## Spectrogram video

You can also create a spectrogram video like this:

```{r}
av_spectrogram_video('short.mp3', output = 'spectrogram.mp4', width = 1280, height = 720, res = 144)
```

<video width="100%" controls>
<source src="spectrogram.mp4" type="video/mp4">
Your browser does not support the video tag.
</video>

## Compare with tuneR/signal

We convert the audio sample to `wav` so that we can read it with tuneR:

```{r}
# Conver to wav audio with av
av_audio_convert('short.mp3', 'short.wav', channels = 1)

# Read wav with tuneR
data <- tuneR::readWave('short.wav')

# demean to remove DC offset
snd <- data@left - mean(data@left)
```

We then use the signal package to calculate the spectrogram with similar parameters as av:

```{r}
# create spectrogram
spec <- signal::specgram(x = snd, n = 1024, Fs = data@samp.rate, overlap = 1024 * 0.75)

# normalize and rescale to dB
P <- abs(spec$S)
P <- P/max(P)
P <- 10*log10(P)

# plot spectrogram
image(x = spec$t, y = spec$f,z = t(P), ylab = 'Freq [Hz]', xlab = 'Time [s]', useRaster=TRUE)
```

## Compare with seewave

Compare spectrograms using the `tico` audio sample included with the seewave package:

```{r}
library(seewave)
library(ggplot2)
data(tico)
ggspectro(tico, ovlp = 50) + geom_tile(aes(fill = amplitude)) + stat_contour()
```

To use av, we first save the wav file and then create spectrogram:

```{r}
# First export wav file
savewav(tico, filename = 'tico.wav')
plot(read_audio_fft('tico.wav'))
```

## Compare with phonTools

Use the audio sample included with phonTools:

```{r}
library(phonTools)
data(sound)
spectrogram(sound, maxfreq = sound$fs/2)
```


Save the wav file and then create spectrogram. We match the default window function from phonTools:

```{r}
phonTools::writesound(sound, 'sound.wav')
plot(read_audio_fft('sound.wav', window = phonTools::windowfunc(1024, 'kaiser')))
```
